Note on the word2vec and the gravity model

### 2023-08-24 paper/current/nature-comm-physics/main.tex
many small changes

\begin{document} %{{{
\title{Unsupervised embedding of trajectories captures the latent structure of discrete mobility.} %{{{
\author[1,\Yinyang]{Dakota Murray}
\author[1,2,\Yinyang]{Jisung Yoon}
\author[1]{Sadamori Kojaku}
\author[3,4]{Rodrigo Costas}
\author[2,5,6]{Woo-Sung Jung}
\author[1]{Staša Milojević \thanks{smilojev@indiana.edu}}
\author[1,7,8]{Yong-Yeol Ahn \thanks{yyahn@iu.edu}}

\affil[1]{Center for Complex Networks and Systems Research, Luddy School of Informatics, Computing, and Engineering, Indiana University, Bloomington, IN 47408, USA}
\affil[2]{Department of Industrial and Management Engineering, Pohang University of Science and Technology,
	Pohang 37673, Republic of Korea.}
\affil[3]{Centre for Science and Technology Studies (CWTS), Leiden University, P.O. Box 905,
	2300 AX Leiden, The Netherlands}
\affil[4]{DST-NRF Centre of Excellence in Scientometrics and Science, Technology and Innovation Policy, Stellenbosch University, Stellenbosch, South Africa}
\affil[5]{Department of Physics, Pohang University of Science and Technology,
	Pohang 37673, Republic of Korea.}
\affil[6]{Asia Pacific Center for Theoretical Physics, Pohang 37673, Republic of Korea.}
\affil[7]{Network Science Institute, Indiana University, Bloomington (IUNI), IN 47408, USA}
\affil[8]{Connection Science, Massachusetts Institute of Technology, Cambridge, MA 02139, USA}

\affil[\Yinyang]{These authors contributed equally: Jisung Yoon, Dakota Murray.}

@@

%
% ABSTRACT
%
\newpage
\begin{sciabstract}
	Human mobility drives major societal phenomena including epidemics, economies, and innovation.
	Historically, mobility was constrained by geographic distance, however, in the globalizing world, language, culture, and history are increasingly important.
	Here, we show a mathematical equivalence between \textit{word2vec} model and the gravity model of mobility and demonstrate that, by using three datasets of human mobility trajectories, \textit{word2vec} encodes nuanced relationships between discrete locations into an accurate and meaningful vector-space, providing a functional distance between locations, as well as a representation for studying the many dimensions of mobility.
	Focusing on the case of scientific mobility, we show that embeddings can learn implicit information directly from mobility data, such as cultural, linguistic, and hierarchical relationships at multiple levels of granularity.
	Connecting neural embeddings to the gravity model opens up new avenues for the study of mobility.
\end{sciabstract}

@@

Here, we demonstrate that the \textit{word2vec} model (Skip-Gram Negative Sampling) is equivalent to the gravity law of mobility, a fundamental mobility framework.
We then empirically test the resulting representation by its ability to derive from mobility trajectories the functional distances between locations.
After validating that it accurately represents real-world data, we leverage unique embedding-based methods to demonstrate \textit{word2vec}'s capacity for encoding rich information relating to geography, culture, language, and even prestige, at multiple scales of analysis.
We embed trajectories from three datasets: U.S. passenger flight itinerary records, Korean accommodation reservations, and a dataset of scientists' career mobility between organizations captured in bibliometric records (Detailed descriptions are available in the Methods).

@@

\begin{equation}
	\label{eq:gravity_basic}
	\hat{T}_{ij} = Cm_{i}m_{j}f(r_{ij}),
\end{equation}
where $m_{i}$ is the population of location $i$,  $f(r_{ij})$ is a decay function with respect to distance between locations, and $C$ is a constant estimated from data (see Methods).
For the flight itinerary data, we use population $m_{i}$ as the total number of unique passengers who passed through each airport, for the Korean accommodation reservation data, we use the total number of unique customers who booked with each accommodation, and for scientific mobility, we use the mean annual number of unique mobile and non-mobile authors who were affiliated with each organization.
$\hat{T}_{ij}$, which is often referred to as ``expected flux''~\autocite{simini2012universal}, is the expected frequency of the co-occurrence of location $i$ and $j$ in the trajectory in the gravity model.

@@

NCE is a generic estimator for probability model~\autocite{Gutmann2010}
\begin{align}
	\label{eq:nce-model}
	P_m(x) = \frac{ f(x) }{\sum_{ x' \in {\cal X}} f(x') },
\end{align}
where $f$ is a positive real-valued likelihood function of data $x$, and ${\cal X}$ is the set of all data.
Note that word2vec belongs to this class of probability models, with $x=\vect{u}_i\cdot \vect{v}_j$ and $f(x)=\exp(x)$.
To train {\it word2vec} with NCE~\autocite{mikolov2013word2vec}, one samples a center-context pair $(i,j)$ from the given data and labels the pair as $Y=1$. 
Then, one replaces the context location $j$ with a random location $j'$ sampled from a noise distribution $p_0(j')$ and labels the pair as $Y=0$. 
NCE finds the embedding that can classify the center-context pairs using a logistic function 
\begin{align}
	\label{eq:nce}
	P^{\text{NCE}}\left(Y_{j}=1 \vert j\right) = \frac{
		1
	}{
		1 + \exp\left[ - \ln f(\vect{u}_j \cdot \vect{v}_{i})  + \ln p_0(j) \right]
	},
\end{align}
by maximizing the log-likelihood
\begin{align}
	\label{eq:log-likelihood-logistic-regress}
	{\cal J}^{\text{NCE}} = \sum_{i \in {\cal A}}\sum_{j \in {\cal D}} \left[ Y_{j} \log P^{\text{NCE}}(Y_{j} = 1\vert j) + (1-Y_{j}) \log P^{\text{NCE}}(Y_{j} = 0 \vert j)\right].
\end{align}
Note that NCE is an unbiased estimator that has asymptomatic convergence to the optimal embedding in terms of the original word2vec's objective function, ${\cal J}$~\autocite{Chia2010,Dyer2014}.

>>>>

\begin{document} %{{{
\title{Unsupervised embedding of trajectories captures the latent structure of discrete mobility.} %{{{
\author[1,\Yinyang]{Dakota Murray}
\author[1,2,3,\Yinyang]{Jisung Yoon}
\author[1, \Yinyang]{Sadamori Kojaku}
\author[4,5]{Rodrigo Costas}
\author[6,7]{Woo-Sung Jung}
\author[1]{Staša Milojević \thanks{smilojev@indiana.edu}}
\author[1,8,9]{Yong-Yeol Ahn \thanks{yyahn@iu.edu}}

\affil[1]{Center for Complex Networks and Systems Research, Luddy School of Informatics, Computing, and Engineering, Indiana University, Bloomington, IN 47408, USA}
\affil[2]{Kellogg School of Management at Northwestern University, Evanston, IL 60208, USA.}
\affil[3]{Northwestern Institute on Complex Systems, Evanston, IL 60208, USA.}
\affil[4]{Centre for Science and Technology Studies (CWTS), Leiden University, P.O. Box 905,
	2300 AX Leiden, The Netherlands}
\affil[5]{DST-NRF Centre of Excellence in Scientometrics and Science, Technology and Innovation Policy, Stellenbosch University, Stellenbosch, South Africa}
\affil[6]{Department of Physics, Pohang University of Science and Technology,
	Pohang 37673, Republic of Korea.}
\affil[7]{Department of Industrial and Management Engineering, Pohang University of Science and Technology,
	Pohang 37673, Republic of Korea.}
\affil[8]{Network Science Institute, Indiana University, Bloomington (IUNI), IN 47408, USA}
\affil[9]{Connection Science, Massachusetts Institute of Technology, Cambridge, MA 02139, USA}

\affil[\Yinyang]{These authors contributed equally: Sadamori Kojaku, Jisung Yoon, Dakota Murray, }

@@

%
% ABSTRACT
%
\newpage
\begin{sciabstract}
	Human mobility drives major societal phenomena including epidemics, economies, and innovation.
	Historically, mobility was constrained by geographic distance, however, in the globalizing world, language, culture, and history are increasingly important.
	Here, we show a mathematical equivalence between \textit{word2vec} model and the gravity model of mobility and demonstrate that, by using three datasets of human mobility trajectories, \textit{word2vec} encodes nuanced relationships between discrete locations into an accurate and meaningful vector-space, providing a functional distance between locations, as well as a representation for studying the many dimensions of mobility.
	Focusing on the case of scientific mobility, we show that embeddings can learn implicit information,  such as cultural, linguistic, and hierarchical relationships at multiple levels of granularity,  directly from mobility data.
	Connecting neural embeddings to the gravity model opens up new avenues for the study of mobility.
\end{sciabstract}

@@

Here, we demonstrate that the \textit{word2vec} model (Skip-Gram Negative Sampling) is equivalent to the gravity law of mobility, a fundamental mobility framework.
We then empirically test the resulting representation by its ability to derive the functional distances between locations from mobility trajectories.
After validating that it accurately represents real-world data, we leverage unique embedding-based methods to demonstrate \textit{word2vec}'s capacity for encoding rich information related to geography, culture, language, and even prestige, at multiple scales of analysis.
We embed trajectories from three datasets: U.S. passenger flight itinerary records, Korean accommodation reservations, and a dataset of scientists' career mobility between organizations captured in bibliometric records (Detailed descriptions are available in the Methods).

@@

\begin{equation}
	\label{eq:gravity_basic}
	\hat{T}_{ij} = Cm_{i}m_{j}f(r_{ij}),
\end{equation}
where $m_{i}$ is the population of location $i$,  $f(r_{ij})$ is a decay function with respect to distance between locations, and $C$ is a constant estimated from data (see Methods).
For the flight itinerary data, we use population $m_{i}$ as the total number of unique passengers who passed through each airport, for the Korean accommodation reservation data, we use the total number of unique customers who booked with each accommodation, and for scientific mobility, we use the mean annual number of unique mobile and non-mobile authors who were affiliated with each organization.
$\hat{T}_{ij}$, which is often referred to as ``expected flux''~\autocite{simini2012universal},  as the expected frequency of the co-occurrence of location $i$ and $j$ in the trajectory in the gravity model.

@@

NCE is a generic estimator for probability model~\autocite{Gutmann2010}
\begin{align}
	\label{eq:nce-model}
	P_m(x) = \frac{ f(x) }{\sum_{ x' \in {\cal X}} f(x') },
\end{align}
where $f$ is a positive real-valued likelihood function of data $x$, and ${\cal X}$ is the set of all data.
Note that {\it word2vec} belongs to this class of probability models, with $x=\vect{u}_i\cdot \vect{v}_j$ and $f(x)=\exp(x)$.
To train {\it word2vec} with NCE~\autocite{mikolov2013word2vec}, one samples a center-context pair $(i,j)$ from the given data and labels the pair as $Y=1$. 
Then, one replaces the context location $j$ with a random location $j'$ sampled from a noise distribution $p_0(j')$ and labels the pair as $Y=0$. 
NCE finds the embedding that can classify the center-context pairs using a logistic function 
\begin{align}
	\label{eq:nce}
	P^{\text{NCE}}\left(Y_{j}=1 \vert j\right) = \frac{
		1
	}{
		1 + \exp\left[ - \ln f(\vect{u}_j \cdot \vect{v}_{i})  + \ln p_0(j) \right]
	},
\end{align}
by maximizing the log-likelihood
\begin{align}
	\label{eq:log-likelihood-logistic-regress}
	{\cal J}^{\text{NCE}} = \sum_{i \in {\cal A}}\sum_{j \in {\cal D}} \left[ Y_{j} \log P^{\text{NCE}}(Y_{j} = 1\vert j) + (1-Y_{j}) \log P^{\text{NCE}}(Y_{j} = 0 \vert j)\right].
\end{align}
Note that NCE is an unbiased estimator that has asymptomatic convergence to the optimal embedding in terms of the original word2vec's objective function, ${\cal J}$~\autocite{Chia2010,Dyer2014}.

### 2023-08-24 paper/current/nature-comm-physics/notes/2021-0105-word2vec-vs-gravity-model/main.tex
many small changes

Now, let us move our focus to the \textit{word2vec}.
In out setting, the model takes a sequence of locations and models the co-occurrence of center location $i$ and context location $j$ by
\begin{equation}
    p(j \given i) = \frac{\exp(\bm{u}_j \cdot \bm{v}_{i})}{Z_i},
\end{equation}
where $\bm{v}$ and $\bm{u}$ are the ``in-vector" and ``out-vector", respectively, and $Z_i=\sum_{x \in \mathcal{A}} \exp(\bm{u}_{x} \cdot \bm{v}_{i})$ is a normalization constant,
and $\mathcal{A}$ is the set of all locations.
Under restricting window size to be one, the expected flow between center location $a_i$ and context location $a_j$ is given by
\begin{align}
    \label{eq:flow_w2v}
    \hat{T_{ij}}=s_i P(j\given i) = \frac{s_i\exp(\bm{u}_j \cdot \bm{v}_{i})}{Z_i},
\end{align}
where $s_i$ is the frequency of location $i$ summed for all trajectories in the dataset.
Although there are some similarities, Eq.~\ref{eq:flow_w2v} does not have a population term $m_i$. In additionm the distance can be asymmetric, $r_{ij}=- \bm{u}_{j} \cdot \bm{v}_{i} \neq - \bm{u}_{i} \cdot \bm{v}_{j}=r_{ji}$.

@@

Parameter $\gamma=1$ is a special choice that ensures that the in-vector and out-vector are the same, \ie $\bm{u}_i = \bm{v}_i$ \autocite{levy2014neural}.
Setting $\gamma = 1$, and substituting $\bm{u}_i = \bm{v}_j$ and $p_0(j) = s_j / \sum_{\ell'} s_{\ell'}$ into Eq~\eqref{eq:flow_w2v_ng0}, we have
\begin{align}
    P\left(j \given i \right) & := \frac{ p_0 (j) \exp(\bm{v}_j \cdot \bm{v}_{i})}{ Z'_i}                      \\
                              & =  \frac{s _j  \exp(\bm{v}_j \cdot \bm{v}_{i})}{ Z' _i\sum_{\ell'} s_{\ell'}},
\end{align}
which leads to the expected flow for the word2vec model, \ie
\begin{align}
    \hat{T_{ij}}=s_i P(j\given i) =  \frac{s_i s _j  \exp(\bm{v}_j \cdot \bm{v}_{i})}{  Z'_i\sum_{\ell'} s_{\ell'}}. \label{eq:flow_w2v_ng0}
\end{align}
In the \textit{word2vec}, the flow $\hat{T_{ij}}_{ij}$ is symmetric by design (\ie $\hat{T_{ij}}=\hat{T_{ji}}$) because the word2vec neglects whether the context $j$ appears later or earlier from center $i$.
Therefore, we have
\begin{align}
    T_{ij} = T_{ji}
     & \Rightarrow
    \frac{s_i s _j  \exp(\bm{v}_j \cdot \bm{v}_{i})}{  Z'_i\sum_{\ell'} s_{\ell'}} =  \frac{s_j s _i  \exp(\bm{v}_i \cdot \bm{v}_{j})}{  Z'_j\sum_{\ell'} s_{\ell'}} \\
     & \Rightarrow
    Z'_i = Z'_j, \label{eq:euqal_rel}
\end{align}
suggesting that the normalization constant $Z' _i$ is constant.
Combining Eqs.~\eqref{eq:flow_w2v_ng0} and \eqref{eq:euqal_rel}, we have
\begin{align}
    \hat{T_{ij}}=  C s_i s _j  \exp(\bm{v}_j \cdot \bm{v}_{i}),
\end{align}
where $C$ is a positive constant.
Eq.~\eqref{eq:flow_w2v_ng0} reveals the connection between the \textit{word2vec} and the gravity model, with with $d_{ij} = - \bm{u}_{j} \cdot \bm{v}_{i}$ and $f(x) = \exp(-x)$.
In sum, the \textit{word2vec} is a gravity model, in which the mass of location $i$ is its frequency in the trajectory data, and the distance between locations is measured by their dot similarities.

@@

Many probability models including the word2vec can be represented in form
\begin{align}
    f(x) / \sum_{x' \in {\cal X}} f(x'),
\end{align}
where $f$ is a nonnegative function.
Fitting the probability models are computationally expensive as its denominator involves a sum over all possible entities ${\cal X}$ in data.
The negative sampling is introduced as a way to avoid computing the denominator.
To this end, the negative sampling solves a synthetic binary classification task, in which
we are asked to classify an actual data $x$ and a noise data $\tilde x$ using a sigmoid function
\begin{align}
    \label{eq:sigmoid}
    \frac{1}{1 + \exp\left( -\log f(z) \right)}.
\end{align}
Because this sigmoid function does not involve the normalization constant, the model fitting can be done with less computational burden.

@@

If the bias is not corrected, what model does the negative sampling actually fit?
To see this, we rewrite the sigmoid function (Eq.~\eqref{eq:sigmoid}) in form of the bias-corrected sigmoid function, \ie
\begin{align}
    \label{eq:sigmoid3}
    \frac{1}{1 + \exp\left[ - \log \tilde f(z) + \log p_0(z)  \right]},
\end{align}
where we have defined
\begin{align}
    \label{eq:unnormalized}
    \tilde f(z) = f(z) p_0(z).
\end{align}
This sigmoid function correctly fits a probability model:
\begin{align}
    f'(x) / \sum_{x' \in {\cal X}} f'(x),
\end{align}
or, by substituting Eq.~\eqref{eq:unnormalized},
\begin{align}
    p_0(x)f(x) / \sum_{x' \in {\cal X}} p_0(x')f(x').
\end{align}
For the word2vec, the correct expression of word2vec optimized with the negative sampling is given by
\begin{align}
    P\left(j \given i \right):= \frac{p_0(j)\exp(\bm{u}_j \cdot \bm{v}_{i})}{Z'_i}.
\end{align}

>>>>

Now, let us move our focus to the \textit{word2vec}.
In out setting, the model takes a sequence of locations and models the co-occurrence of center location $i$ and context location $j$ by
\begin{equation}
    p(j \given i) = \frac{\exp(\bm{u}_j \cdot \bm{v}_{i})}{Z_i},
\end{equation}
where $\bm{v}$ and $\bm{u}$ are the ``in-vector" and ``out-vector", respectively,  $Z_i=\sum_{x \in \mathcal{A}} \exp(\bm{u}_{x} \cdot \bm{v}_{i})$ is a normalization constant, and $\mathcal{A}$ is the set of all locations.
Under restricting window size to be one, the expected flow between center location $i$ and context location $j$ is given by
\begin{align}
    \label{eq:flow_w2v}
    \hat{T_{ij}}=s_i P(j\given i) = \frac{s_i\exp(\bm{u}_j \cdot \bm{v}_{i})}{Z_i},
\end{align}
where $s_i$ is the frequency of location $i$ summed for all trajectories in the dataset.
Although there are some similarities, Eq.~\ref{eq:flow_w2v} does not have a population term $m_i$. In addition, the distance can be asymmetric, $r_{ij}=- \bm{u}_{j} \cdot \bm{v}_{i} \neq - \bm{u}_{i} \cdot \bm{v}_{j}=r_{ji}$.

@@

Parameter $\gamma=1$ is a special choice that ensures that the in-vector and out-vector are the same, \ie $\bm{u}_i = \bm{v}_i$ \autocite{levy2014neural}.
Setting $\gamma = 1$, and substituting $\bm{u}_i = \bm{v}_j$ and $p_0(j) = s_j / \sum_{\ell'} s_{\ell'}$ into Eq~\eqref{eq:flow_w2v_ng0}, we have
\begin{align}
    P\left(j \given i \right) & := \frac{ p_0 (j) \exp(\bm{v}_j \cdot \bm{v}_{i})}{ Z'_i}                      \\
                              & =  \frac{s _j  \exp(\bm{v}_j \cdot \bm{v}_{i})}{ Z' _i\sum_{\ell'} s_{\ell'}},
\end{align}
which leads to the expected flow for the word2vec model, \ie
\begin{align}
    \hat{T_{ij}}=s_i P(j\given i) =  \frac{s_i s _j  \exp(\bm{v}_j \cdot \bm{v}_{i})}{  Z'_i\sum_{\ell'} s_{\ell'}}. \label{eq:flow_w2v_ng0}
\end{align}
In the \textit{word2vec}, the flow $\hat{T_{ij}}_{ij}$ is symmetric by design (\ie $\hat{T_{ij}}=\hat{T_{ji}}$) because the model neglects whether the context $j$ appears later or earlier from center $i$.
Therefore, we have
\begin{align}
    T_{ij} = T_{ji}
     & \Rightarrow
    \frac{s_i s _j  \exp(\bm{v}_j \cdot \bm{v}_{i})}{  Z'_i\sum_{\ell'} s_{\ell'}} =  \frac{s_j s _i  \exp(\bm{v}_i \cdot \bm{v}_{j})}{  Z'_j\sum_{\ell'} s_{\ell'}} \\
     & \Rightarrow
    Z'_i = Z'_j, \label{eq:euqal_rel}
\end{align}
suggesting that the normalization constant $Z' _i$ is constant.
Combining Eqs.~\eqref{eq:flow_w2v_ng0} and \eqref{eq:euqal_rel}, we have
\begin{align}
    \hat{T_{ij}}=  C s_i s _j  \exp(\bm{v}_j \cdot \bm{v}_{i}),
\end{align}
where $C$ is a positive constant.
Eq.~\eqref{eq:flow_w2v_ng0} reveals the connection between the \textit{word2vec} and the gravity model, with with $d_{ij} = - \bm{u}_{j} \cdot \bm{v}_{i}$ and $f(x) = \exp(-x)$.
In sum, the \textit{word2vec} is a gravity model, in which the mass of location $i$ is its frequency in the trajectory data, and the distance between locations is measured by their dot similarities.

@@

Many probability models including the \textit{word2vec} can be represented in form
\begin{align}
    f(x) / \sum_{x' \in {\cal X}} f(x'),
\end{align}
where $f$ is a nonnegative function.
Fitting the probability models are computationally expensive as its denominator involves a sum over all possible entities ${\cal X}$ in data.
The negative sampling is introduced as a way to avoid computing the denominator.
To this end, the negative sampling solves a synthetic binary classification task, in which
we are asked to classify an actual data $x$ and a noise data $\tilde x$ using a sigmoid function
\begin{align}
    \label{eq:sigmoid}
    \frac{1}{1 + \exp\left( -\log f(z) \right)}.
\end{align}
Because this sigmoid function does not involve the normalization constant, the model fitting can be done with less computational burden.

@@

If the bias is not corrected, what model does the negative sampling actually fit?
To see this, we rewrite the sigmoid function (Eq.~\eqref{eq:sigmoid}) in form of the bias-corrected sigmoid function, \ie
\begin{align}
    \label{eq:sigmoid3}
    \frac{1}{1 + \exp\left[ - \log \tilde f(z) + \log p_0(z)  \right]},
\end{align}
where we have defined
\begin{align}
    \label{eq:unnormalized}
    \tilde f(z) = f(z) p_0(z).
\end{align}
This sigmoid function correctly fits a probability model:
\begin{align}
    f'(x) / \sum_{x' \in {\cal X}} f'(x),
\end{align}
or, by substituting Eq.~\eqref{eq:unnormalized},
\begin{align}
    p_0(x)f(x) / \sum_{x' \in {\cal X}} p_0(x')f(x').
\end{align}
For the \textit{word2vec}, the correct expression optimized with the negative sampling is given by
\begin{align}
    P\left(j \given i \right):= \frac{p_0(j)\exp(\bm{u}_j \cdot \bm{v}_{i})}{Z'_i}.
\end{align}

### 2023-08-24 paper/current/nature-comm-physics/notes/gm-vs-w2v/main.tex
many small changes

% for quick author comments 
\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
\definecolor{light-gray}{gray}{0.8}
\definecolor{light-blue}{RGB}{33,150,243}
\definecolor{light-green}{RGB}{76, 175, 80}
\def\del#1{ {\color{light-gray}{#1}} }
\def\yy#1{\footnote{\color{red}\textbf{yy: #1}} }
\def\sada#1{\footnote{\color{light-blue}\textbf{sk: #1}} }
\def\jisung#1{\footnote{\color{light-green}\textbf{js: #1}} }

@@

With the word2vec, one writes the probability of target node $j$ given center node $i$ as 
\begin{align}
  \label{eq:model_node2vec}
    P_{\text{W2V}}\left(j \given i\right) := \frac{1}{Z_i}\exp\left(\vec{u} ^\top _{i} \vec{v}_{j}\right), 
\end{align}
where vectors $\vec{u}_i$ and $\vec{v}_i$ are the column vectors of length $K$ that represent the coordinates of $i$ and $j$ in the embedding space spanned by in-vectors and out-vectors, respectively.
Variable $Z_i$ is the normalization constant, i.e., $Z_i:= \sum_{\ell=1}^N \exp\left(\vec{u} ^\top _{i} \vec{v}_{\ell}\right)$. 
Denote by $d(\vec{u},\vec{u}')$ the distance between the in-vectors, i.e., 
\begin{align}
    \label{eq:distance}
    d(\vec{u},\vec{u}'):= \sqrt{\sum_{k=1}^K \left(u_{k} - u'_{k}\right)^2}.
\end{align}
We assume that the in-vectors are equal to the out-vectors, i.e., $\vec{v}_j = \vec{u}_j$.
Then, the dot similarity for $\vec{u}_i$ and $\vec{v}_j$ is given by 
\begin{align}
    \label{eq:dotsim}
    \vec{u}_i ^\top \vec{v}_j = \frac{1}{2}\left[ ||\vec{u}_i||^2 + ||\vec{u}_j||^2 - d\left(\vec{u}_i, \vec{u}_j\right)^2 \right],
\end{align}
which leads 
\begin{align}
    \exp\left( \vec{u}_i ^\top \vec{v}_j \right) &= \frac{\exp\left( ||\vec{u}_i||^2/2 + ||\vec{u}_j||^2/2 \right)}{\exp\left[ d\left(\vec{u}_i, \vec{u}_j\right)^2 / 2 \right]} \nonumber \\
                                                 &= \frac{s_i s_j}{\exp\left[d\left(\vec{u}_i, \vec{u}_j\right)^2 / 2\right]}, \label{eq:exp_distance_form}
\end{align}
where $s_i:=\exp(||\vec{u} _i||^2/2)$. 
Substituting Eq.~\eqref{eq:exp_distance_form} into Eq.~\eqref{eq:model_node2vec} yields
\begin{align}
    \label{eq:model_node2vec_distance}
    P_{\text{W2V}}\left(j \given i\right) &:= \frac{1}{Z_i}\cdot \frac{s_is_j}{\exp\left[d\left(\vec{u}_i, \vec{u}_j\right)^2/2\right]}. 
\end{align}
The walkers in point $i$ tend to move to the points with a large $s_j$ more than those with a smaller $s_j$.
Thus, variable $s_j$ represents the {\it pulling force} for node $j$\sada{tentative naming}. 
It turns out that $s_j$ is a mass variable for the gravity model (see Section 2).

@@

In migration study, $M_i$ is often fixed to the number of people at a point $i$ (e.g., city and country), which is a restricted version of the gravity model.
We show that the word2vec is free from the restriction and automatically learns the mass $M_i$ from data, which may not be equal to the number of people at a point.

@@

What is pulling force $s_i$? Is it related to the mobile population $\pi_i$ in each point?
To answer the questions, we rewrite $\pi_i$ as a function of $s_i$ as follows. 
With Eq.~\eqref{eq:const}, we rewrite $\pi_i$ as 
\begin{align}
    \pi_i &= G' Z_i \nonumber \\
        &= G' s_i \sum_{\ell=1}^N s_\ell f\left(d\left(\vec{u}_i,\vec{u}_\ell\right)\right) \nonumber \\ 
        &= G' s_i ^2 - s_i \Phi(\vec{u}_i), \label{eq:flow_gravity_model_2} %\nonumber \\ 
\end{align}
where $\Phi(\vec{u})$ is the potential given by 
\begin{align}
    \Phi(\vec{u}):=-G' \sum_{\ell=1, \vec{u}\neq \vec{u}_\ell}^N s_\ell f\left(d\left(\vec{u},\vec{u}_\ell\right)\right).
\end{align}
Potential $\Phi(\vec{u})$ is analogous to the gravitational potential, i.e., $-GM/d$, where $M$ is the mass and $d$ is the distance from the mass.
In other words,  $\Phi(\vec{u})$ is the gravitational potential for the field, where 
each point $\ell$ has a mass $s_{\ell}$ and is away from $\vec{u}$ with distance $d(\vec{u},\vec{u}_\ell)$, and the effect of the mass decays by the factor of $\exp\left[d(\vec{u},\vec{u}_\ell) ^2 /2\right]$.

>>>>

% comment out when submitting
% for quick author comments 
\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
\usepackage{soul}
\usepackage{ulem}
\definecolor{light-gray}{gray}{0.8}
\definecolor{light-blue}{RGB}{33,150,243}
\definecolor{light-green}{RGB}{76, 175, 80}
%\def\del#1{ {\color{light-gray}{#1}} }
\def\yy#1{\footnote{\color{red}\textbf{yy: #1}} }
\def\sada#1{\footnote{\color{light-blue}\textbf{sk: #1}} }
\def\jisung#1{\footnote{\color{light-green}\textbf{js: #1}} }

% Track Changes
\definecolor{myblue}{RGB}{33,150,243}
\definecolor{mygreen}{RGB}{76, 175, 80}
\definecolor{purple}{RGB}{170, 0, 255}

\newcommand{\add}[1]{{\leavevmode\color{blue}#1}}
\newcommand{\del}[1]{{\leavevmode\color{red}\sout{#1}}}
\newcommand{\change}[3][]{{\del{#2}\add{#3}$^{\textcolor{blue}{\text{#1}}}$}}
\newcommand{\comment}[1]{
    \begin{enumerate}[label*=(\arabic*)]
    #1
    \end{enumerate}
}

@@

With the word2vec, one writes the probability of target node $j$ given center node $i$ as 
\begin{align}
  \label{eq:model_node2vec}
    P_{\text{W2V}}\left(j \given i\right) := \frac{1}{Z_i}\exp\left(\vec{u} ^\top _{i} \vec{v}_{j}\right), 
\end{align}
where vectors $\vec{u}_i$ and $\vec{v}_i$ are the column vectors of length $K$ that represent the coordinates of $i$ and $j$ in the embedding space spanned by in-vectors and out-vectors, respectively.
Variable $Z_i$ is the normalization constant, i.e., $Z_i:= \sum_{\ell=1}^N \exp\left(\vec{u} ^\top _{i} \vec{v}_{\ell}\right)$. 
Denote by $d(\vec{u},\vec{u}')$ the distance between the in-vectors, i.e., 
\begin{align}
    \label{eq:distance}
    d(\vec{u},\vec{u}'):= \sqrt{\sum_{k=1}^K \left(u_{k} - u'_{k}\right)^2}.
\end{align}
We assume that the in-vectors are equal to the out-vectors, i.e., $\vec{v}_j = \vec{u}_j$.
Exploiting $d^2 (\vec{u},\vec{u}')=||\vec{u}||^2 + ||\vec{u}'||^2 - 2\vec{u} ^\top \vec{u}'$, the dot similarity for $\vec{u}_i$ and $\vec{v}_j$ is given by 
\begin{align}
    \label{eq:dotsim}
    \vec{u}_i ^\top \vec{v}_j = \frac{1}{2}\left[ ||\vec{u}_i||^2 + ||\vec{u}_j||^2 - d\left(\vec{u}_i, \vec{u}_j\right)^2 \right],
\end{align}
which leads 
\begin{align}
    \exp\left( \vec{u}_i ^\top \vec{v}_j \right) &= \frac{\exp\left( ||\vec{u}_i||^2/2 + ||\vec{u}_j||^2/2 \right)}{\exp\left[ d\left(\vec{u}_i, \vec{u}_j\right)^2 / 2 \right]} \nonumber \\
                                                 &= \frac{s_i s_j}{\exp\left[d\left(\vec{u}_i, \vec{u}_j\right)^2 / 2\right]}, \label{eq:exp_distance_form}
\end{align}
where $s_i:=\exp(||\vec{u} _i||^2/2)$. 
Substituting Eq.~\eqref{eq:exp_distance_form} into Eq.~\eqref{eq:model_node2vec} yields
\begin{align}
    \label{eq:model_node2vec_distance}
    P_{\text{W2V}}\left(j \given i\right) &:= \frac{1}{Z_i}\cdot \frac{s_is_j}{\exp\left[d\left(\vec{u}_i, \vec{u}_j\right)^2/2\right]}. 
\end{align}
The walkers in point $i$ tend to move to the points with a large $s_j$ more than those with a smaller $s_j$.
Thus, variable $s_j$ represents the {\it pulling force} for node $j$\sada{tentative naming}. 
It turns out that $s_j$ is a mass variable for the gravity model (see Section 2).

@@

How should $M_i$ be chosen? The choice of $M_i$ is entirely unconstrained.
In migration study, one sets $M_i$ to exogenous variables such as population at the point in priori and then estimates other parameters, e.g., $G$.
We show that the word2vec automatically estimates the mass $M_i$ from data without exogenous variables. 
The estimated $M_i$ may not be equal to the number of people at a point.

@@

What is pulling force $s_i$? Is it related to the mobile population $\pi_i$ in each point?
To answer the questions, we rewrite $\pi_i$ as a function of $s_i$ as follows.
With Eq.~\eqref{eq:const}, we rewrite $\pi_i$ as 
\begin{align}
    \pi_i &= G' Z_i \nonumber \\
        &= G' s_i \sum_{\ell=1}^N s_\ell f\left(d\left(\vec{u}_i,\vec{u}_\ell\right)\right)\\
        &= -G' s_i \Phi(\vec{u}_i), \label{eq:flow_gravity_model_2} \\ %\nonumber \\ 
\end{align}
where we have defined $\Phi(\vec{u})$ by 
\begin{align}
    \Phi(\vec{u}):=-G' \sum_{\ell=1}^N s_\ell f\left(d\left(\vec{u},\vec{u}_\ell\right)\right).
\end{align}
In other words, $\pi_i$ is proportional to the product of $s_i$ and $\Phi(\vec{u})$.
Therefore, points with a large $\pi_i$ tend to have a large $s_i$ value.
What does $\Phi(\vec{u})$ represent? 
Variable $\Phi(\vec{u})$ is analogous to the gravitational potential in mechanical physics, which takes form $-GM/d$, where $M$ is the mass and $d$ is the distance from the mass.
%In other words,  $\Phi(\vec{u})$ can be seen as the gravitational potential for the field, where 
%each point $\ell$ has a mass $s_{\ell}$ and is away from $\vec{u}$ with distance $d(\vec{u},\vec{u}_\ell)$, and the effect of the mass decays by the factor of $\exp\left[d(\vec{u},\vec{u}_\ell) ^2 /2\right]$.
Point $\vec{u}$ has a large $\Phi(\vec{u})$ if there are many points $\ell$ around $\vec{u}$ that have a large $\vec{s}_\ell$.
Potential $\Phi(\vec{u})$ is similar to the eigen centrality for networks that assigns a large centrality score for the node with many neighbors with a large centrality scores.
